---
title: "Deep 7 FRS Weight Comp Investigation"
author: "Meg Oshima"
date: "7/26/2021"
output: 
  html_document:
    toc: true
    toc_float: true

---


```{r setup, include=FALSE}
library(tidyverse)
library(gt)
library(nmfspalette)
library(sf)
library(ggridges)

knitr::opts_chunk$set(warning=FALSE,
                message=FALSE,
                echo=FALSE)  

scale_colour_continuous <- function(...) scale_fill_nmfs(palette = "regional web")
scale_colour_discrete   <- function(...) scale_color_nmfs(palette = "regional web")
scale_fill_continuous <- function(...) scale_fill_nmfs(palette = "regional web")
scale_fill_binned <- function(...) scale_color_nmfs(palette = "regional web")


HDAR_areas <- st_read("../Data/GIS/Shapefiles/DAR_Reporting_grids_all.shp")
mhi_cropped <- st_crop(HDAR_areas, xmin = -162, xmax = -153,
                          ymin = 17, ymax = 24)

deep7_frs <- read.csv("../Data/Deep_7_FRS.csv")

```


### Full Weight Data Set  
Data are FRS records for single catch fish with weights up to the state record weights for each species. Data were collected from 1948 to 2019. Only deep-sea handline gear type was used because it represents between 94 to 98% of the total pounds caught.  

```{r}


com_names <- c("Hapuupuu","Kalekale","Opakapaka","Ehu","Onaga", "Lehi","Gindai")
st_records <- data.frame(COMMON_NAME = com_names,  RECORD = c(70,4,19,12,35,33,5))

st_records %>% 
  arrange(desc(RECORD)) %>% 
  gt() %>% 
  cols_label(
    COMMON_NAME = "Common Name",
    RECORD = "State Record (lbs)"
  ) %>% 
  tab_header(title = "State Records for Deep 7 Species")

```

<br>
```{r}
deep7_frs %>% 
  ggplot(aes(x = COMMON_NAME, y = LBS)) +
  geom_boxplot() +
  facet_wrap(~COMMON_NAME, scales = "free") +
  theme_classic() +
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

### Weight Compositions for Each Species  

```{r}
deep7_frs %>% 
  ggplot(aes(x = LBS)) +
  geom_density() +
  facet_wrap(~COMMON_NAME, scales = "free") +
  theme_classic() 
```

```{r}

deep7_frs %>% 
  mutate(Decade = floor(FYEAR / 10) * 10,
         Decade = factor(Decade, levels = seq(1940, 2010, by = 10))) %>% 
  ggplot(aes(x = LBS, group = Decade, color = Decade)) +
  geom_density() +
  facet_wrap(~COMMON_NAME, scales = "free") +
  theme_classic() +
  labs(title = "Weight Distributions by Decade")
```

### Weight Compositions by Area Type

```{r}
Ehu.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Ehu"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Ehu")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Ehu.type, aes(x = 10, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Ehu")

```


```{r}
Gin.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Gindai"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Gindai")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Gin.type, aes(x = 4, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Gindai")

```


```{r}
Hapu.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Hapuupuu"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Hapuupuu")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Hapu.type, aes(x = 300, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Hapuupuu")

```



```{r}
Kale.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Kalekale"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Kalekale")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Kale.type, aes(x = 3, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Kalekale")

```


```{r}
Lehi.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Lehi"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Lehi")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Lehi.type, aes(x = 20, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Lehi")

```


```{r}
Ona.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Onaga"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Onaga")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Ona.type, aes(x = 20, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Onaga")

```

```{r}
Opa.type <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(TYPE, COMMON_NAME) %>% 
  rename(Type = "TYPE") %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Opakapaka"))

deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>%
  mutate(Type = factor(TYPE)) %>% 
  group_by(Type, COMMON_NAME) %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(Type, "MHI")) %>% 
  filter(str_detect(COMMON_NAME, "Opakapaka")) %>% 
  ggplot(aes(x = LBS, y = Type, fill = Type)) +
  geom_density_ridges(alpha = .5) +
  geom_text(data = Opa.type, aes(x = 15, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Opakapaka")

```

For Ehu, Gindai, Hapuupuu, and Kalekale, there is not much difference in weight composition by area type. For Lehi, Onaga, and Opakapaka, inshore and coastal areas have very similar distributions but offshore shows different distribution. More, larger fishes.  

### Weight Compositions by Island  


```{r}

ehu.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Ehu"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Ehu")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = ehu.island, aes(x = 10, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Ehu")


```


```{r}

gin.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Gindai"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Gindai")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = gin.island, aes(x = 4, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Gindai")


```



```{r}

hapu.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Hapuupuu"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Hapuupuu")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = hapu.island, aes(x = 300, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Hapuupuu")


```


```{r}

kale.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Kalekale"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Kalekale")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = kale.island, aes(x = 3.5, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Kalekale")


```


```{r}

lehi.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Lehi"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Lehi")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = lehi.island, aes(x = 30, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Lehi")


```


```{r}

ona.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Onaga"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Onaga")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = ona.island, aes(x = 22, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Onaga")


```



```{r}

opa.island <- deep7_frs %>% 
  merge(mhi_cropped, by.x = "AREA", by.y = "AREA_ID", all = TRUE) %>% 
  filter(!is.na(LBS)) %>%
  group_by(Region, COMMON_NAME) %>% 
  summarise(N = paste0("N = ", n())) %>% 
  filter(str_detect(COMMON_NAME, "Opakapaka"))

deep7_frs %>% 
  filter(!is.na(LBS)) %>%
  filter(str_detect(COMMON_NAME, "Opakapaka")) %>% 
  ggplot(aes(x = LBS, y = Region, fill = Region)) +
  geom_density_ridges(alpha = .5, show.legend = FALSE) +
  geom_text(data = opa.island, aes(x = 17, y = , label = N), position = position_nudge(y = +.5)) +
  theme_classic() +
  scale_fill_nmfs("regional web") +
  labs(title = "Opakapaka")


```

Lehi and Opakapaka are the only 2 species that have some regional differences in weight compositions. Niihau distributions are the most different from the other islands for both species.


### Converting Weights to Length Comps  


Using the equation $$L = {(W/\alpha)}^{1/\beta}$$ where:  

  *   L = Length (cm)       
  *   W = Weight (kg)       
  *   $\alpha$ = scaling parameter (from Ault et al. 2018)        
  *   $\beta$ = volumetric parameter (from Ault et al. 2018)    

we converted weights to lengths and look at the distribution. 


```{r}
# convert from weight to length by L = (W/alpha)^(1/beta)

camera_lengths <- read.csv("../Data/BFISH 2016-2019 Camera Lengths.csv")
fishing_lengths <- read.csv("../Data/BFISH 2016-2019 Research Fishing Lengths.csv")

deep_7 <- unique(camera_lengths$COMMON_NAME)[-c(8:9)]


cam_7 <- camera_lengths %>% 
  filter(str_detect(BFISH, "_S", negate = TRUE)) %>% 
  filter(COMMON_NAME %in% deep_7) %>% 
  separate(BFISH, into = c("BFISH", "Year", "Seas"), sep = "_") %>% 
  rename(SAMPLE_ID = "DROP_CD",
         DEPTH = OFFICIAL_DEPTH_M) %>% 
  select(PSU, 
         SAMPLE_ID, 
         Year, 
         SPECIES_CD, 
         COMMON_NAME, 
         DEPTH, 
         LENGTH_CM, 
         Island) %>% 
  mutate(Gear = "Camera")

fish_7 <- fishing_lengths %>% 
  filter(str_detect(BFISH, "_S", negate = TRUE)) %>% 
  filter(COMMON_NAME %in% deep_7) %>% 
  separate(BFISH, into = c("BFISH", "Year", "Seas"), sep = "_") %>% 
  rename(DEPTH = SAMPLE_MEAN_DEPTH_M) %>% 
  select(PSU, 
         SAMPLE_ID,
         Year, 
         SPECIES_CD, 
         COMMON_NAME, 
         DEPTH, 
         LENGTH_CM, 
         Island) %>% 
  mutate(Gear = "Research Fishing")

wl.params <- data.frame(
  COMMON_NAME = c("Ehu", "Gindai", "Hapuupuu", "Kalekale", "Lehi", "Onaga", "Opakapaka"),
  alpha = c(1.551E-05, 3.526E-05, 3.065E-05, 2.243E-05, 1.298E-04, 6.005E-05, 2.311E-05),
  beta = c(3.026, 2.859, 2.884, 2.932, 2.458, 2.673, 2.928))

deep7_frs %>% 
  mutate(kg = (LBS *0.45359237)) %>% 
  merge(wl.params, by = "COMMON_NAME") %>% 
  mutate(Length = (kg/alpha)^(1/beta)) %>% 
  ggplot(aes(x = Length)) +
  geom_density(colour = "#C95E15") +
  geom_density(data = cam_7, aes(x = LENGTH_CM), color = "#00467F") +
  geom_density(data = fish_7, aes(x = LENGTH_CM), color = "#157253") +
  facet_wrap(~COMMON_NAME, scales = "free") +
  theme_classic() +
  labs(caption = "Blue line represents the BFISH camera survey data, green line represents the BFISH research fishing data, \n and orange line represents the converted lengths from FRS weights data.")
  
```

